Ab Initio Effective One-Electron Potentials Revisited: General Theory and Example Applications for Charge-Transfer Energy Calculations

### 2019-05-21 papers/2_oep_technique/paper-r0.tex
Paper II CT-OEP with added discussion on OEP basicts from CI expansion

\title{Ab Initio Effective One-Electron Potentials 
for Repulsive and Charge-Transfer Interaction Energies at Hartree-Fock Level}

@@

\begin{abstract}
The concept of the effective one-electron potentials has been useful for many decades
in efficient description of electronic structure of chemical systems, especially extended
molecular aggregates such as interacting molecules in condensed phases. 
In this Work, \emph{ab initio} effective potentials are studied for a number of theories
of the intemolecular interaction energy.
\end{abstract}

@@

The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. Formally, the effective density can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation}
	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation}
	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
	\ket{\alpha} 
	\left[
	\int
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
	\rho^{\rm eff}({\bf r})
	d{\bf r} 
	\right]
	\bra{\beta}  
\end{equation}
%
with the Coulomb operator defined by
%
\begin{equation}
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
	f({\bf r}) 
	\int
	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%
for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tbraket{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \sum_{\gamma\delta} \braket{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
\end{equation}
%
In the above equation and throughout the work, the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\braket{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

@@

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem,
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $,
$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%
\begin{equation}
	\BraKet{ij}{kl} \equiv
	\iint 
	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;,
\end{equation}
%
whereas the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
%In this work we are interested in cases when

@@

Application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} and adopting the localized basis set approximation
yields the partitioning of the interaction\hyp{}induced property,
%
\begin{equation}
 \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
\end{equation}
%
for which the one\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-1el}
 \Delta^{AB} P(1) \approx
	\sum_{ij} \Big\{ 
	C^{AA}_{ij}  \BraKet{A}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
	C^{BB}_{ij}  \BraKet{B}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	C^{AB}_{ij}  \BraKet{A}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{multline}
%
and the two\hyp{}electron part is given by
%
\begin{multline} \label{e:delta-p-part-2el}
 \Delta^{AB} P(2) \approx
	\sum_{ijkl} \Big\{ 
	C^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
	C^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
	{s} \left[ C^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\} \;.
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\BraKet{A}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.
All the summation components from Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el} 
can be grouped into the three
sub\hyp{}categories: (i) Coulomb\hyp{}like terms
with integrals of the type $\BraKet{AA}{BB}$, (ii) overlap\hyp{}like terms with $\BraKet{AA}{AB}$
and $\BraKet{AB}{BB}$ and (iii) exchange\hyp{}like terms with $\BraKet{AB}{AB}$.

@@

Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
ERI's. 
%
 \begin{equation}
 \mathcal{F}
 \left[ 
   \ERIc{i}{j}{k^B}{l^B}
	 \right] = \OEIc{i}{v_{kl}^B}{j}  \;,
 \end{equation}
%
where 
% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
%or basis function.
$v_{kl}^B$ is the effective one-electron potential derived from the partial
effective densities $\rho_{kl}^B({\bf r}) = \phi_k^B({\bf r})\phi_l^B({\bf r})$.
The summations over $k$ and $l$ can be incorporated into the generalized one-electron potential
$v_{\text{eff}}^B$
to produce
%
\begin{equation}
	\sum_{ij}\sum_{kl\in B} \mathcal{F}\left[ 
   \ERIc{i}{j}{k^B}{l^B}
 \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j}  \;.
\end{equation}
%
Thus, the total computational effort is, in principle, reduced from the fourth-fold
sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
It is also possible to generalize the above expression even further by
summing over all possible functionals ${\mathcal{F}}_t$
%
\begin{equation} \label{e:ft-reduction}
	\sum_t \sum_{ij}\sum_{kl\in B} {\mathcal{F}}_t\left[ 
   \ERIc{i}{j}{k^B}{l^B}
 \right] = \sum_{ij} \OEIc{i}{v_{\text{eff}}^B}{j} \;.
\end{equation}
%
The above design has the advantage that the effective potentials are fully first-principles
and no extensive case-dependent fitting procedures are necessary.
Only the effective potentials of \emph{independent} fragment $B$ need to be determined once and for all
and stored in a file. Note also, that, in principle, there is no approximation 
made here at that moment.

@@

The above technique can be now applied to the Coulomb\hyp{}like and overlap\hyp{}like 
terms from Eq.~\eqref{e:delta-p-part-2el}. For example, the Coulomb\hyp{}like contributions
can be rewritten without making any approximation as
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ C^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
 \equiv
	\sum_{kl\in B} 
	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ C^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
\end{equation}
%
Similarly for the overlap\hyp{}like term we have
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ C^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
 \equiv
	\sum_{k\in A} \sum_{l\in B} 
	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
and the associated effective potential reads
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ C^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
\end{equation}
%
Note also that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
%
\begin{multline}
 \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
	+ \sum_{i\in A} \sum_{j\in B} 
	\left\{
	  \tBraKet{i}{\hat{v}^{{\rm eff},A}_1 }{j}
	+ \tBraKet{i}{\hat{v}^{{\rm eff},B}_1 }{j}
	\right\} \\ +
	\frac{1}{2}
	\sum_{i\in A} \sum_{k\in A}
	  \tBraKet{i}{\hat{v}^{{\rm eff},B}_2 }{k}
	 +
	 \frac{1}{2}
        \sum_{l\in B} \sum_{l\in B}
          \tBraKet{l}{\hat{v}^{{\rm eff},A}_2 }{l}
\end{multline}
%
where the exchange\hyp{}like component is
%
\begin{equation}
	\Delta_{\rm ex}^{AB} P = 
	\sum_{ik\in A} \sum_{jl\in B}
	{s} \left[ C^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] \;.
\end{equation}
%
whereas
the overlap\hyp{}like and the Coulomb\hyp{}like 
OEP's are defined as follows
%
\begin{align}
	\hat{v}^{{\rm eff},A}_1 &\equiv C_{ij}^{AB} 
	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
	%
	\hat{v}^{{\rm eff},A}_2 &\equiv 
	+ \sum_{kk'\in A} {s} \left[ C^{AAAB}_{kk'ij} \hat{v}_{kk'}^{AA} \right] \\
\end{align}
%
and accordingly for the twin terms (molecule $B$).

\section{\label{s.334}Practical Calculations with OEP's}

There are two general cases of basis function partitioning scheme
in which it is possible to define an effective one\hyp{}electron
potential.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, can be approximated with the density fitting. Mainly,
%
\begin{subequations}
 \begin{align}
 \mathcal{F}
 \left[ 
   \ERIc{i^A}{j^A}{k^B}{l^B}
	 \right] = \OEIc{i^A}{v_{kl}^B}{j^A} \\ \nonumber \rightarrow \text{ multipole expansion or density fitting}\\ 
%
 \mathcal{F}
 \left[ 
   \ERIc{i^A}{j^B}{k^B}{l^B}
	 \right] = \OEIc{i^A}{v_{kl}^B}{j^B} \\ \rightarrow \text{ density fitting} \nonumber
 \end{align}
\end{subequations}
%

>>>>

\title{Ab Initio Effective One-Electron Potentials Revisited:
General Theory and Example Applications for Charge-Transfer Energy Calculations}

@@

\begin{abstract}
The concept of the effective one-electron potentials has been useful for many decades
in efficient description of electronic structure of chemical systems, especially extended
molecular aggregates such as interacting molecules in condensed phases. 
%In this Work, \emph{ab initio} effective potentials are studied for a number of theories
%of the intemolecular interaction energy.
\end{abstract}

@@

The one\hyp{}electron potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. Formally, the effective density can be expanded in terms of an effective
one\hyp{}particle density matrix represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}({\bf r}) = \sum_{\alpha\beta} D_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = \sum_{\alpha\beta} 
	\Ket{\alpha} 
	\left[
	\int
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) 
	\rho^{\rm eff}({\bf r})
	d{\bf r} 
	\right]
	\Bra{\beta}  
\end{equation}
%
with the Coulomb operator defined by
%
\begin{equation}
	\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) \equiv
	f({\bf r}) 
	\int
	\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%
for any one\hyp{}electron function $f({\bf r})$. The matrix element of the effective potential operator
is therefore given by
%
\begin{equation}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} D^{\rm eff}_{\gamma\delta}  \;.
\end{equation}
%
In the above equation and throughout the work, 
%
\begin{equation}
\tBraKet{\alpha}{\mathcal{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathcal{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathcal{O}(1)$ 
and the electron repulsion integral (ERI)
is defined according to
%
\begin{equation}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

@@

The time\hyp{}independent Schr{\"o}dinger equation provides the stationary solution to the
electronic state
of a molecular system in terms of its $N$\hyp{}electron wavefunction, $\Psi$,
%
\begin{equation} \label{e:schrodinger}
 E = \tBraKet{\Psi}{\mathscr{H}}{\Psi} \;,
\end{equation}
%
for a quantum Hamiltonian, $\mathscr{H} = \hat{h}(1) + \hat{o}(2)$,
with $\hat{h}(1)$ being the core Hamiltonian one\hyp{}electron operator
and $\hat{o}(2)$ the two\hyp{}electron repulsion operator.
From the above solution, any property $P$, understood as an expectation value of the
associated operator $\mathscr{P}$, can be computed according to
%
\begin{equation} \label{e:prop}
 P = \tBraKet{\Psi}{\mathscr{P}}{\Psi} \;.
\end{equation}
%
The exact solution to Eq.~\eqref{e:schrodinger} is formally given by
the configuration interaction (CI) expansion around the reference 
(approximate) wavefunction, $\Psi^{(0)}$,
typically the solution to the Hartree\hyp{}Fock (HF) problem,
%
\begin{equation} \label{e:ci}
 \Ket{\Psi} = \Ket{\Psi^{(0)}} + \sum_{ra} C_{r}^{a} \Ket{\Psi_{r}^{a}} + 
	 \sum_{rsab} C_{rs}^{ab} \Ket{\Psi_{rs}^{ab}} + \ldots
\end{equation}
%
In the above equation,
$\Ket{\Psi_{rs\ldots}^{ab\ldots}}$ are called the CI configurations
with the associated CI coefficients $C_{rs\ldots}^{ab\ldots}$, created by `exciting'
one or more electrons from the $a$th ($b$th) occupied molecular orbital to the $a$th ($b$th)
virtual (unoccupied) molecular orbital, obtained from the HF solution.
By inserting Eq.~\eqref{e:ci} into Eq.~\eqref{e:prop} 
and using the Slater\hyp{}Condon rules for evaluating the one\hyp{} and two\hyp{}electron
operator matrix element between
the CI configurations, it can be shown that the following holds
%
\begin{equation} \label{e:exp-val-series}
	P =
	\sum_{ij} P_{ij} \tBraKet{i}{\hat{h}}{j}
	+ \sum_{ijkl} P_{ijkl} \BraKet{ij}{kl}  \;,
\end{equation}
%
%where $\tBraKet{i}{\hat{h}}{j} \equiv \int \phi^*_i({\bf r}) \hat{h} \phi_j({\bf r}) d{\bf r} $
%$\BraKet{ij}{kl}$ is the electron\hyp{}repulsion integral (ERI) defined by
%%
%\begin{equation}
%	\BraKet{ij}{kl} \equiv
%	\iint 
%	\frac{ \phi_i^{*}({\bf r}_1) \phi_j({\bf r}_1) 
%	       \phi_k^{*}({\bf r}_2) \phi_l({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
%	d{\bf r}_1 d{\bf r}_2  \;,
%\end{equation}
%%
where the second\hyp{}rank tensor ${\bf P}^{(2)}$ 
and the fourth\hyp{}rank tensor ${\bf P}^{(4)}$ 
are well defined and characteristic for the property of interest.
In fact, they are certain functions of the CI coefficients, i.e.,
%
\begin{equation}
{\bf P} = {\bf P}
\left(
 \left\{
  C_{r}^{a}, C_{rs}^{ab}, \ldots 
 \right\} 
\right)
\end{equation}
%
and, in principle, are known as long as the solution to the CI problem 
or its approximation is available.

%In this work we are interested in cases when

@@

Application of Eq.~\eqref{e:exp-val-series}
to Eq.~\eqref{e:delta-p} and adopting the localized basis set approximation
yields the partitioning of the interaction\hyp{}induced property,
%
\begin{equation} \label{e:delta-p-part}
 \Delta^{AB} P = \Delta^{AB} P(1) + \Delta^{AB} P(2) \;,
\end{equation}
%
for which the one\hyp{}electron part reads
%
\begin{multline} \label{e:delta-p-part-1el}
 \Delta^{AB} P(1) \approx
	\sum_{ij} \Big\{ 
	P^{AA}_{ij}  \tBraKet{A}{\hat{h}}{A} +    %\tBraKet{i^A}{\hat{h}}{j^A} + 
	P^{BB}_{ij}  \tBraKet{B}{\hat{h}}{B} \\ + %\tBraKet{i^B}{\hat{h}}{j^B} + 2 
2	P^{AB}_{ij}  \tBraKet{A}{\hat{h}}{B}      %\tBraKet{i^A}{\hat{h}}{j^B} 
	\Big\}
\end{multline}
%
and the two\hyp{}electron part is given by
%
\begin{multline} \label{e:delta-p-part-2el}
 \Delta^{AB} P(2) \approx
	\sum_{ijkl} \Big\{ 
	P^{AAAA}_{ijkl} \BraKet{AA}{AA} \\ +    %\BraKet{i^Aj^A}{k^Al^A} +
	P^{BBBB}_{ijkl} \BraKet{BB}{BB} +  %\BraKet{i^Bj^B}{k^Bl^B} +
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right] \\ + 
	{s} \left[ P^{ABAB}_{ijkl} \BraKet{AB}{AB} \right] + 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right] \\ +
	{s} \left[ P^{ABBB}_{ijkl} \BraKet{AB}{BB} \right]
	\Big\} \;.
\end{multline}
%
To simplify the resulting expressions
in Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el},
the shorthand notation $\tBraKet{A}{\hat{h}}{B} \equiv \tBraKet{i^A}{\hat{h}}{j^B}$ and 
$\BraKet{AB}{CD} \equiv \BraKet{i^Aj^B}{k^Cl^D}$
was introduced, and
the real orbitals 
were assumed to simplify the symmetry of the one\hyp{} and two\hyp{}electron integrals.
The operator $s$ permutes and sums over all the combinations of the basis function indices
given their partitioning scheme between molecules $A$ and $B$.
All the summation components from Eqs.~\eqref{e:delta-p-part-1el} and \eqref{e:delta-p-part-2el} 
can be grouped into the three
sub\hyp{}categories: (i) Coulomb\hyp{}like terms
with integrals of the type $\BraKet{AA}{BB}$, (ii) overlap\hyp{}like terms with $\BraKet{AA}{AB}$
and $\BraKet{AB}{BB}$ and (iii) exchange\hyp{}like terms with $\BraKet{AB}{AB}$.

@@

Consider now an arbirtary functional $\mathcal{F}$ that explicitly depends on the 
ERI's. In this work, OEP's are defined by the following transformation
%
 \begin{equation}
 \mathcal{F}
 \left[ 
   \BraKet{ij}{k^Al^A}
	 \right] = \tBraKet{i}{\hat{v}_{kl}^A}{j}  \;,
 \end{equation}
%
where 
% $A$ and $B$ denote different molecules and $\phi_i$ is the $i$th molecular orbital
%or basis function.
$\hat{v}_{kl}^A$ is the effective one-electron potential operator given by Eq.~\eqref{e:oep-operator} 
with the
effective density $\rho_{kl}^A({\bf r}) \equiv \phi_k^A({\bf r})\phi_l^A({\bf r})$.
The summations over $k$ and $l$ can be incorporated into the total effective one-electron potential operator
$\hat{v}_{\text{eff}}^A$
to produce
%
\begin{equation}
	\sum_{ij}\sum_{kl\in A} \mathcal{F}\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j}  \;.
\end{equation}
%
Thus, the total computational effort is, in principle, reduced from the fourth-fold
sum involving evaluation of ERI's to the two-fold sums of cheaper one-electron integrals.
It is also possible to generalize the above expression even further by
summing over all possible functionals ${\mathcal{F}}_t$
%
\begin{equation} \label{e:ft-reduction}
	\sum_t \sum_{ij}\sum_{kl\in A} {\mathcal{F}}_t\left[ 
   \BraKet{ij}{k^Al^A}
 \right] = \sum_{ij} \tBraKet{i}{\hat{v}_{\text{eff}}^A}{j} \;.
\end{equation}
%
The above design has the advantage that it opens the possibility to define first\hyp{}principles
effective fragments as long as the $P_{ij}$ and $P_{ijkl}$ 
from Eq.~\eqref{e:exp-val-series} are computable and can be approximately
partitioned in between the interacting fragments.
%the resulting effective potentials are fully first\hyp{}principles
%and no extensive case\hyp{}dependent fitting procedures are necessary as long as the $P_{ij}$ and $P_{ijkl}$
%are computable.
%Only the effective potentials of \emph{independent} fragment $A$ need to be determined once and for all
%and stored in a file. 
%Note also, that, in principle, there is no approximation 
%made here at that moment.

@@

The above technique can be now applied to the CI expansion of the interaction\hyp{}induced property
under the localized basis set approximation 
from Eq.~\eqref{e:delta-p-part}.
%Coulomb\hyp{}like and overlap\hyp{}like 
%terms from Eq.~\eqref{e:delta-p-part-2el}. 
For example, the Coulomb\hyp{}like contributions
can be rewritten without making any further approximation as
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AABB}_{ijkl} \BraKet{AA}{BB} \right]
 \equiv
	\sum_{kl\in B} 
	\tBraKet{k^B}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
with
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AABB}_{ijkl} \hat{v}^{AA}_{ij} \right]  \;.
\end{equation}
%
Similarly for the overlap\hyp{}like term we have
%
\begin{equation}
 \sum_{ijkl} 
	{s} \left[ P^{AAAB}_{ijkl} \BraKet{AA}{AB} \right]
 \equiv
	\sum_{k\in A} \sum_{l\in B} 
	\tBraKet{k^A}{\hat{v}^{{\rm eff},A}}{l^B}
\end{equation}
%
and the associated effective potential reads
%
\begin{equation}
 \hat{v}^{{\rm eff},A} 
	\equiv \sum_{ij\in A} {s} \left[ P^{AAAB}_{ijkl} \hat{v}^{AA}_{ij} \right] \;.
\end{equation}
%
Note that the exchange\hyp{}like terms cannot be represented in terms of OEP's.
Gathering the above arguments, $\Delta^{AB}P$ adopts the following form
%
\begin{multline}
 \Delta^{AB}P \approx \Delta_{\rm ex}^{AB} P
	+ \sum_{i\in A} \sum_{j\in B} 
	  \tBraKet{i}{ \left\{ \hat{v}^{{\rm eff},AB}_{1(ij)} + \hat{v}^{{\rm eff},BA}_{1(ij)}  \right\} }{j}
	 \\ +
	\sum_{i\in A} \sum_{k\in A}
	  \tBraKet{i}{\hat{v}^{{\rm eff},BA}_{2(ik)} }{k}
	 +
        \sum_{j\in B} \sum_{l\in B}
          \tBraKet{j}{\hat{v}^{{\rm eff},AB}_{2(jl)} }{l}
\end{multline}
%
where the exchange\hyp{}like component is
%
\begin{equation}
	\Delta_{\rm ex}^{AB} P = 
	s \sum_{ik\in A} \sum_{jl\in B}
	P^{ABAB}_{ijkl} \BraKet{ij}{kl} \;,
\end{equation}
%
and
the overlap\hyp{}like and the Coulomb\hyp{}like 
OEP's are defined as follows
%
\begin{align}
	\hat{v}^{{\rm eff},BA}_{1(ij)} &\equiv P_{ij}^{AB} \hat{h}
	+ s \sum_{mk\in A}  P^{AAAB}_{mkij} \hat{v}_{mk}^{AA} \\
	%
	\hat{v}^{{\rm eff},BA}_{2(ik)} &\equiv P_{ik}^{AA} \hat{h} 
        + s \sum_{mk'\in A} P^{AAAA}_{ikmk'} \hat{v}_{mk'}^{AA}  \nonumber
        \\
	&+ \frac{1}{2}
        s \sum_{jl'\in B}  P^{AABB}_{ik'jl} \hat{v}_{jl'}^{BB} 
\end{align}
%
and accordingly for the twin operators $\hat{v}^{{\rm eff},AB}_{\text{$1$ and $2$}}$.

\section{\label{s.334}Practical OEP-Based Calculations}

There are two general cases of basis function partitioning scheme
in which it is possible to define an effective one\hyp{}electron
potential. They are listed in Table~\ref{t:oep-matrix-element-types}.
The first type is a Coulomb\hyp{}like contribution, for which multipole expansion or density fitting can be used.
The second type is an overlap\hyp{}like contribution, can be approximated with the density fitting. Mainly,
%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{{\rm eff},A}}{j} $ && $\tBraKet{j}{\hat{v}^{{\rm eff},A}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
Associated ERI's    &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &   \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%
